---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r global_options, include=FALSE}
library(knitr)
opts_chunk$set(fig.path='figure/',
                warning=FALSE, message=FALSE,echo=TRUE)
```

## Loading and preprocessing the data

```{r }
# unzip the activity.zip file into the current directory
unzip("./activity.zip",exdir="./")

# read the activity.csv file in and define the column classes
activitydata <- read.csv(file = "activity.csv", na.strings="NA",colClasses=c("integer","POSIXct","integer"))

# create a clean data set with the NA's ommitted
activitydata.clean <- na.omit(activitydata)
```

## What is mean total number of steps taken per day?

-  Make a histogram of the total number of steps taken each day

```{r}
# create an array which holds the sum of steps per day
temp<-tapply(activitydata.clean$steps,format(activitydata.clean$date,"%Y-%m-%d"),sum)

# draw the historgram and set the labes for axis and title
hist(temp,xlab="Sum steps per day",ylab="Frequency", col = "light blue",labels=TRUE,main="Histogram of the total number of steps taken each day")

```

- Calculate and report the mean and median total number of steps taken per day

Take the variable temp that was created in the previous section which has the total steps per day.  Use the is.na to look for any NAs to remove and then calculate the mean and median.


```{r }
# calculate the mean and store in variable for reporting and then also display
steps.mean <- mean(temp[!is.na(temp)])
print(steps.mean)
  
# calculate the median and store in variable for reporting and then also display
steps.median <- median(temp[!is.na(temp)])
print(steps.median)

```



## What is the average daily activity pattern?

- Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

Using the aggregate function we are able to group the 5 min intervals over the days and compute the average.  This is stored in a new variable for plotting



```{r}
aveDailyPattern <- aggregate(steps ~ interval, activitydata.clean, mean)
plot(aveDailyPattern,type="l",ylab="Average steps")
```


- Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r}
# use which.max() to get the row number with the max steps and use this to retrieve the interval from the vector
aveDailyPattern[which.max(aveDailyPattern$steps),1]
```


## Inputing missing values

- Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
sum(!complete.cases(activitydata))

```

- Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

- Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
#

```

- Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}

```

## Are there differences in activity patterns 
//between weekdays and weekends?

- Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}

```

- Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). The plot should look something like the following, which was creating using simulated data:
```{r}

```

